---
title: "Prepare selected time-series"
output: html_document
date: "2023-12-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F)
```

### Goal

Because of the many complications from our different deployment strategies, we're only using a small chunk of the full dataset. Specifically, we want to use data from after the pumped heads were installed (XX/XX) and before one of the pumped heads failed (XX/XX).

First, let's bring in our CO2 data and identify the datetimes where we want our datasets to start and end.


```{r Setup}

## Location of raw EXO datasets
csense_filepath = "../data/csense/timeseries"

## Load constants
source("../scripts/0_constants.R")

```


```{r Import raw csense}

## Set threshold for flagging CO2 based on standard deviation
co2_sd_threshold = 50 #if the sd > threshold, cull

## Helper function
mean_ <- function(var){mean({{var}}, na.rm = T)}
sd_ <- function(var){sd({{var}}, na.rm = T)}

csense_files <- list.files(csense_filepath, pattern = "CR6Series_Table1", full.names = T)

read_cr6 <- function(file){
  df_raw <- read_delim(file, skip = 1) %>% 
    slice(3:n()) %>% 
     mutate(SEVolt_1 = as.numeric(SEVolt_1), 
            SEVolt_2 = as.numeric(SEVolt_2), 
            SEVoltage_1 = as.numeric(SEVoltage_1), 
            SEVoltage_2 = as.numeric(SEVoltage_2)) %>% 
     mutate(datetime_1min = force_tz(parsedate::parse_date(TIMESTAMP), tzone = common_tz)) %>% 
     mutate(datetime = round_date(datetime_1min, unit = time_interval)) %>% 
     mutate(datetime_raw = as.character(datetime)) %>% 
     #select(datetime, datetime_raw, contains("SEVolt_"))  %>% 
     group_by(datetime) %>% 
     summarize(co2_ppm_bare = mean_(SEVolt_1), 
               co2_ppm_eelgrass = mean_(SEVolt_2),
               sd_ppm_bare = sd_(SEVolt_1), 
               sd_ppm_eelgrass = sd_(SEVolt_2), 
               co2_mv_bare = mean_(SEVoltage_1), 
               co2_mv_eelgrass = mean_(SEVoltage_2),
               sd_mv_bare = sd_(SEVoltage_1), 
               sd_mv_eelgrass = sd_(SEVoltage_2)) 
}

csense_raw <- csense_files %>% 
  map(read_cr6) %>% 
  bind_rows()

csense_start <- "2023-07-28"
csense_end <- "2023-09-02"

csense_trim_start <- csense_raw %>% 
  filter(datetime >= csense_start)

csense_trim <- csense_trim_start %>% 
  filter(datetime <= csense_end)
```

### Raw C-sense data {.tabset}

#### All data

```{r}
csense_raw %>% 
  select(datetime, contains("co2_ppm")) %>% 
  pivot_longer(cols = -c(contains("datetime"))) %>% 
ggplot(aes(datetime, value, color = name)) + 
  geom_line()
```

#### Period of interest

Based on plotly, things go off the rails for the eelgrass tank sometime on 9/2, so we'll use that to cut the end.

```{r}
csense_trim_start %>% 
  select(datetime, contains("co2_ppm")) %>% 
  pivot_longer(cols = -c(contains("datetime"))) %>% 
ggplot(aes(datetime, value, color = name)) + 
  geom_line()
```

#### Final cut

```{r}
csense_trim %>% 
  select(datetime, contains("co2_ppm")) %>% 
  pivot_longer(cols = -c(contains("datetime"))) %>% 
ggplot(aes(datetime, value, color = name)) + 
  geom_line()
```

###

### Remaining tasks for CO2

1. Figure out how to correct
2. download final CO2 grab sample data
3. Match with available CO2 grab samples
4. If that's not a lot, take all CO2 grab samples from periods with trustworthy CO2 data


### Pull in EXO data

Based on the timeframe we're using for C-Sense, let's pull in EXO datasets for the relevant time-period

```{r EXO prep}

exo_directory = "../data/exo/timeseries"

## Check file formats (script needs UTF-8) - should be empty!
check_encoding(exo_directory) 

exo_files <- tibble(file = list.files(path = exo_directory, full.names = T)) %>% 
  filter(grepl("230807|230821|230926", file))

```

```{r read EXO data}
df_bare <- exo_files %>% 
  filter(grepl("bare", file)) %>% 
  pull() %>% 
  map(read_exo) %>% 
  bind_rows() %>% 
  rename("p_h1" = p_h_22, "p_h2" = p_h_23)

df_eelgrass <- exo_files %>% 
  filter(grepl("eelgrass", file)) %>% 
  pull() %>% 
  map(read_exo) %>% 
  bind_rows() %>% 
  rename("p_h1" = p_h_21, "p_h2" = p_h_22)

exo <- bind_rows(df_bare %>% mutate(tank = "Bare"), 
                df_eelgrass %>% mutate(tank = "Eelgrass")) 

exo_trim <- exo %>% 
  filter(datetime >= csense_start) %>% 
  filter(datetime <= csense_end)

ggplot(exo_trim, aes(datetime, p_h1, color = tank)) + 
  geom_line()
  

```

```{r}

csense_formatted <- csense_trim %>% 
  select(datetime, contains("co2_ppm")) %>% 
  pivot_longer(cols = -c(contains("datetime"))) %>% 
  mutate(tank = str_to_sentence(str_remove_all(name, "co2_ppm_"))) %>% 
  rename("co2_ppm" = value) %>% 
  select(-name)

df1 <- inner_join(exo_trim, csense_formatted, by = c("datetime", "tank"))
```

```{r}
ggplot(df1, aes(p_h1, co2_ppm, color = tank)) + 
  geom_point()
```


