---
title: "CO2 correlations"
author: "Kay"
date: "2023-07-03"
output: html_document
---

Load libraries
```{r}
require(pacman)
p_load(tidyverse,
       janitor,
       hms,
       dplyr,
       ggpubr, 
       ggplot2,
       corrplot, 
       cowplot,
       readxl,
       data.table,
       parsedate,
       lubridate,
       gridExtra,
       here)
```

Use 'Here' function to map to directory 
```{r}
here::i_am("README.md")
```

Import Excel File of grab sample data. Clean up Data. 
```{r}
biweekly <- read_excel(here("data/Bi-weekly_Tank_Sampling_2023.xlsx"), 2) %>%
  select(contains("Sample ID"), contains("CO2 mean"), contains("Date"), contains("pH (NBS)"), contains("Tank T (C)"), contains ("Salinity")) %>%
  clean_names() %>% slice(4:n())

biweekly <- biweekly %>% drop_na (tank_t_c) 

###biweekly

```


Import Excel File of YSI daily data 
```{r}
### Read in the data from Excel
ysidata <- read_excel(here("data/Daily_TankSpotChecks_2023.xlsx"))

### Select rows and columns, rename columns, select columns for df 
ysidata <- ysidata[2:164,] %>% ### CHANGE THIS AS MORE DATA IS ADDED TO EXCEL 
  rename("sample_id" = "Sample ID", "t_surface" = "T (C)", "t_deep" = "...4", "salinity_surface" = "Salinity (ppt)", "salinity_deep" = "...6", "pH_surface" = "pH (NBS)", "pH_deep" = "...8", "do_surface" = "DO (mg/L)", "do_deep" = "...10", "date"="Date") %>%
  select(-Notes,-Initials,-`Calibration (y/n)`, - `Time Start(PSD)`, -`Time Stop(PSD)`, -`Flow Rate (L/min)`) 

### Convert chars to numeric from data frame, change sample id to eelgrass densities
ysidata <- ysidata %>%
  mutate_at(c('t_surface', 't_deep', 'salinity_surface', 'salinity_deep', 'pH_surface', 'pH_deep', 'do_surface','do_deep'), as.numeric)  

### Select just surface DO (since gas samples taken from surface), date, and sample ID to merge with CO2 dataset 

for_merging <- ysidata %>% select (do_surface, date, sample_id)

###for_merging
```

Merge datasets together 
```{r}
### Merge dataframes together 
merged <- merge (x=biweekly, y=for_merging, by=c('date','sample_id'))

### Some of the columns are still characters, convert them to numeric 
merged <- merged %>%
  mutate_at(c('p_h_nbs', 'tank_t_c', 'salinity_psu'), as.numeric)

merged <- merged %>% 
  mutate(density = case_when(sample_id == 1 ~ 'Medium', sample_id == 2 ~ 'High', sample_id == 3 ~ 'High', sample_id == 4 ~ 'Low', sample_id == 5 ~ 'Zero', sample_id == 6 ~ 'Low', sample_id == 7 ~ 'Zero', sample_id == 8 ~ 'Medium', sample_id == "Inlet" ~ "Inlet"))

merged$density <- factor(merged$density, levels = c("High", "Medium", "Low", "Zero", "Inlet"))

###merged
```

Plot correlation matrices between variables using corrplot
```{r}
co2_cor <- merged %>% 
  select(tank_t_c, p_h_nbs, do_surface, salinity_psu, co2_mean) %>%
  select(where(is.numeric)) %>% ## only select numeric variables
  drop_na()

cor(co2_cor) %>%
  corrplot()


```

CO2 correlations with pH and do
```{r}
### do and co2 correlation stats
summary(lm(do_surface~co2_mean, data = merged))

### pH and co2 correlation stats
summary(lm(do_surface~co2_mean, data=merged))

### salinity and co2 correlation stats 
summary(lm(salinity_psu~co2_mean, data=merged))

### temp and co2 correlation stats
summary(lm(tank_t_c~co2_mean, data=merged))

```

CO2 and Temperature plot 
```{r}
ggplot(na.omit(merged), aes(tank_t_c, co2_mean)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "CO2 versus Temperature", x = "Temperature (C)", y = "CO2 Concentrations (ppm)") +
  theme_set(theme_bw())

```

CO2 and pH plot 
```{r}
ggplot(na.omit(merged), aes(p_h_nbs, co2_mean)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "CO2 versus pH", x = "pH (NBS)", y = "CO2 Concentrations (ppm)") +
  theme_set(theme_bw())

ggsave("co2_ph_lm.png",width = 8, height = 4)

```

Plot co2 and salinity 
```{r}
ggplot(na.omit(merged), aes(salinity_psu, co2_mean)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "CO2 versus Salinity", x = "Salinity (psu)", y = "CO2 Concentrations (ppm)") +
  theme_set(theme_bw())
```

Plot co2 and do 
```{r}
ggplot(na.omit(merged), aes(tank_t_c, co2_mean)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "CO2 versus Dissolved Oxygen", x = "Dissolved Oxygen (mg/L)", y = "CO2 Concentrations (ppm)") +
  theme_set(theme_bw())

ggsave("co2_do_lm.png",width = 8, height = 4)
```

Plot co2 for different densities  
```{r}
compare_density = compare_density <- list(c("Medium", "Low"),
                                          c("Low", "High"),
                                          c("High", "Medium"), 
                                          c("High", "Zero"), 
                                          c("Zero", "Medium"), 
                                          c("Zero","Low"))


ggplot(na.omit(merged), aes(density, co2_mean, fill=density)) + 
  geom_boxplot() + 
  labs(title = "CO2 Concentration in Tanks of Different Densities of Eelgrass", x="Density of Eelgrass", y="CO2 concentration (ppm)") +
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_density)

ggsave("co2_densities_boxplot.png",width = 8, height = 4)


```


Plot CO2 vs DO linear regression with densities separated by color 
```{r}
ggplot(na.omit(merged), aes(do_surface, co2_mean, color=date)) + 
  geom_point() + 
  geom_smooth(method = "lm", se=FALSE) + 
  facet_wrap(~density) +
  labs(title = "CO2 versus Dissolved Oxygen by Density of Eelgrass", x = "Dissolved Oxygen (mg/L)", y = "CO2 Concentrations (ppm)") +
  theme_set(theme_bw())  
  ###scale_color_gradient(low="green", high="black")

ggsave("co2_do_densbycolor.png",width = 8, height = 4)

```

Plot DO and pH over time for different densities. Use ysidata 
```{r}
### Get Data ready 
ysidata <- ysidata %>%
  mutate(density = case_when(sample_id == 2 ~ 'High', sample_id == 3 ~ 'High', sample_id == 1 ~'Medium', sample_id == 8 ~ 'Medium', sample_id == 4 ~ 'Low', sample_id == 6 ~ 'Low', sample_id == 5 ~ 'Zero', sample_id == 7 ~ 'Zero', sample_id == "Inlet" ~ "Inlet"))

ysidata$density <- factor(ysidata$density, levels = c("High", "Medium", "Low", "Zero", "Inlet"))

ysidata %>% 
  filter(!is.na(sample_id)) %>% 
  select(-sample_id) %>% 
  pivot_longer(-date, 
               names_to = c("Var", ".value"), 
               names_sep = "_")


## Create two dataframes, one for surface, one for deep
df_surface <- ysidata %>% 
  filter(!is.na(sample_id)) %>% 
  select(-sample_id) %>%
  select(date, density, contains("_surface")) %>% ## Select just surface columns + metadata
  rename_with(~str_remove(., '_surface')) %>% ## Remove _surface so columns will match
  mutate(depth = "surface")

df_deep <- ysidata %>% 
  filter(!is.na(sample_id)) %>% 
  select(-sample_id) %>%
  select(date, density, contains("_deep")) %>%  
  rename_with(~str_remove(., '_deep')) %>% 
  mutate(depth = "deep")

df <- bind_rows(df_surface, df_deep)

### PLOTS 

### do over time
ggplot(na.omit(df), aes(date, do, color=density)) + 
  geom_point() + 
  geom_smooth(method = "lm", se=FALSE) + 
  ###facet_wrap(~density) +
  labs(title = "Dissolved Oxygen Over Time by Density of Eelgrass", x = "Date", y = "Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())

ggsave("do_time.png",width = 8, height = 4)

### pH over time 
ggplot(na.omit(df), aes(date, pH, color=density)) + 
  geom_point() + 
  geom_smooth(method = "lm", se=FALSE) + 
  facet_wrap(~density) +
  labs(title = "pH Over Time by Density of Eelgrass", x = "Date", y = "pH (NBS)") +
  theme_set(theme_bw())

?geom_smooth

ggsave("ph_time.png",width = 8, height = 4)

```
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
EXO AND CO2 BIWEEKLY+TIMESERIES CORRELATIONS 

Look at EXO data for the tanks compared to the CO2 data and timeseries.

Read in EXO data 
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```{r}
#load in biweekly datasheet 
co2_grab <- read_excel(here("data/Bi-weekly_Tank_Sampling_2023.xlsx"),2) %>% 
  select(contains("Sample ID"), contains("CO2 mean"), contains("CO2 stdev"), contains("Date"), contains("Time Collected (PSD)")) %>%
  clean_names() %>% slice(4:n())
#select only the rows corresponding to tanks 3 and 5
co2_grab <- co2_grab %>% filter(grepl(paste(c(3,5), collapse='|'), sample_id))  %>% drop_na(time_collected_psd)
#for some reason the times in this sheet get loaded in as decimals (ie 0.5 is noon)
#manually convert to seconds and drop anything that doesnt have a co2 measurement
co2_grab <- co2_grab %>% mutate(time_collected_psd = as.hms(as.numeric(time_collected_psd)*3600*24))
#select all the calgas rows and create a datetime column
co2_grab <- co2_grab %>% mutate(datetime = as.POSIXct(paste(parse_date(date), time_collected_psd)))

#load in timeseries sheet
timeseries <- read_excel(here("data/Bi-weekly_Tank_Sampling_2023.xlsx"),4) %>% 
  select(contains("Sample ID"), contains("CO2 mean"), contains("CO2 stdev"), contains("Date"), contains("Time Collected (PSD)")) %>%
  clean_names()
#for some reason the times get saved with a random date attached, reformat to lose the date
timeseries$time_collected_psd <- format(ymd_hms(timeseries$time_collected_psd), "%H:%M:%S")
#select all the calgas rows and create a datetime column
timeseries <- timeseries %>% filter(grepl(paste(c(3,5), collapse='|'), sample_id)) %>%
  mutate(datetime = as.POSIXct(paste(parse_date(date), time_collected_psd))) 


#merges and pivots the grab sample data 
grabs <- rbind(timeseries, co2_grab)

grabs <- grabs %>% select(sample_id, co2_mean, datetime) %>%
  mutate(datetime = force_tz(datetime, tz = "America/Los_Angeles"),
         co2_mean = 1.052*co2_mean)

grabs <- grabs %>% pivot_wider(names_from = sample_id, values_from = co2_mean)
grabs <- grabs %>% rename("co2_5"="5", "co2_3"="3")

View (grabs)

```

Read in EXO data

```{r}
exodata <- read_delim(here("data/exo_timeseries_raw.csv"), show_col_types = FALSE) 
#for pH, take the average of the columns from the 2 pH sensors and create a new column to #store that info in
#make sure the timezone is correct, ie PDT
exodata <- exodata %>% mutate(p_havg = rowMeans(exodata[, c('p_h1', 'p_h2')]),
                              datetime_pdt = force_tz(datetime_pdt, tz = "America/Los_Angeles")) 
#select the relevant columns and pivot so that there are different columns for the #measurements from each tank
exodata <- exodata %>% select(datetime_pdt, temp_c, sal_psu, do_mgl, p_havg, tank)
exodata <- exodata %>% pivot_wider(names_from = tank, names_sep = "_", values_from = c(temp_c, sal_psu, do_mgl, p_havg))

view(exodata)

```

Merge the EXO and CO2 biweekly/timeseries all together 

```{r}
exodata <- data.table(exodata)
co2data <- data.table(grabs)

#sets keys to align time
setkey(co2data, "datetime")
setkey(exodata, "datetime_pdt")

aligned <- exodata[co2data, roll = TRUE]

View (aligned)
```

Plot CO2 versus DO linear regression by tank type: high density or zero density 

```{r}

do_co2_bare <- ggplot(aligned, aes(do_mgl_Bare, co2_5)) + 
  geom_point() + 
  geom_smooth(method = "lm", color="blue") + 
  labs(title = "Zero Density Eelgrass", x="", y="") +
  theme_set(theme_bw()) + theme(plot.title = element_text(hjust = 0.5, size = 12))

do_co2_eelgrass <- ggplot(aligned, aes(do_mgl_Eelgrass, co2_3))+
  geom_point()+
  geom_smooth(method='lm', color="green")+
  labs(title = "High Density Eelgrass", x="", y="") +
  theme_set(theme_bw()) + theme(plot.title = element_text(hjust = 0.5, size = 12))

do_co2 <- grid.arrange(do_co2_bare, do_co2_eelgrass, nrow = 1, top = "CO2 versus DO", 
                       left = "Carbon Dioxide Concentrations (ppm)", 
                       bottom = "Dissolved Oxygen (mg/L)")

ggsave("co2_do.png", do_co2, width=8, height=4)


```


```{r}

ph_co2_bare <- ggplot(aligned, aes(p_havg_Bare, co2_5)) + 
  geom_point() + 
  geom_smooth(method = "lm", color="blue") + 
  labs(title = "Zero Density Eelgrass", x="", y="") +
  theme_set(theme_bw()) + theme(plot.title = element_text(hjust = 0.5, size = 12))

ph_co2_eelgrass <- ggplot(aligned, aes(p_havg_Eelgrass, co2_3))+
  geom_point()+
  geom_smooth(method='lm', color="green")+
  labs(title = "High Density Eelgrass", x="", y="") +
  theme_set(theme_bw()) + theme(plot.title = element_text(hjust = 0.5, size = 12))

ph_co2 <- grid.arrange(ph_co2_bare, ph_co2_eelgrass, nrow = 1, top = "CO2 versus pH", 
                       left = "Carbon Dioxide Concentrations (ppm)", 
                       bottom = "pH")

ggsave("co2_ph.png", ph_co2, width=8, height=4)

```

Look at statistics for pH and CO2, do and CO2 

```{r}
summary(lm(co2_5~p_havg_Bare, data = aligned))

summary (lm(co2_3~p_havg_Eelgrass, data=aligned))

```
pivot longer the data to make box plot by density

```{r}
### rename data to _Bare or _Eelgrass

bydensity <- aligned %>% rename ("co2_Eelgrass"="co2_3", "co2_Bare"="co2_5")

bydensity %>% 
  pivot_longer(-datetime_pdt, 
               names_to = c("Var", ".value"), 
               names_sep = "_")

## Create two dataframes, one for surface, one for deep
df_bare <- bydensity %>% 
  select(datetime_pdt, contains("_Bare")) %>% ## Select just Bare columns
  rename_with(~str_remove(., '_Bare')) %>% ## Remove _surface so columns will match
  mutate(density = "Zero")

df_eelgrass <- bydensity %>% 
  select(datetime_pdt, contains("_Eelgrass")) %>%  
  rename_with(~str_remove(., '_Eelgrass')) %>% 
  mutate(density = "High")

df <- bind_rows(df_bare, df_eelgrass)

view(df)
```


CO2 boxplots to see difference between high and low density eelgrass 

```{r}

compare_density <- list(c("High", "Zero"))

ggplot(df, aes(density, co2, fill = density)) +
  labs(title = "CO2 in Mesocosm Tanks", x = "Eelgrass Density", y = "CO2 Concentrations (ppm)") + 
  geom_boxplot() + ## add boxplots
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_density)

ggsave ("co2_density.png", width = 10, height = 5)

```




