---
title: "Daily.YSI.data"
author: "Kay"
date: "2023-06-15"
output: html_document
---


Use the 'here' function to map to directory 

``` {r}
here::i_am("YSIdata/YSIdata.Rmd") 
```


Load packages

``` {r}
require(pacman)
p_load(tidyverse,
       dplyr,
       ggpubr, 
       ggplot2,
       corrplot, 
       cowplot,
       readxl,
       here)
```


Import Excel file. Clean up data. 

``` {r}
ysidata <- read_excel(here("YSIdata/data/Daily_TankSpotChecks_2023.xlsx"))

ysidata <- ysidata[2:82,] %>% 
  ###mutate (across(where(is_character), as.numeric))
  mutate_if (is_character, as.numeric) %>%
  rename("SampleID" = "Sample ID", "T_Surface" = "T (C)", "T_Deep" = "...4", "Salinity_Surface" = "Salinity (ppt)", "Salinity_Deep" = "...6", "pH_Surface" = "pH (NBS)", "pH_Deep" = "...8", "DO_Surface" = "DO (mg/L)", "DO_Deep" = "...10") %>%
  select(-Notes,-Initials,-`Calibration (y/n)`, - `Time Start(PSD)`, -`Time Stop(PSD)`, -`Flow Rate (L/min)`) 

### not working!!
###ysidata <- mutate_at(c('T_Surface', 'T_Deep', 'Salinity_Surface', 'Salinity_Deep', ###'pH_Surface', 'pH_Deep', 'DO_Surface','DO_Deep'), as.numeric)

###ysidata 

```


Plot the relationship between Surface pH and Surface Temperature in all the tanks. 

``` {r}
ggplot(na.omit(ysidata), aes(T_Surface, pH_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~SampleID) + 
  labs(title = "Surface pH versus Temperature in Mesocosm Tanks", x = "Surface Temperature (C)", y = "Surface pH (NBS)") +
  theme_set(theme_bw())

```
Overall Surface Temp versus pH graph 
```{r}
ggplot(na.omit(ysidata), aes(T_Surface, pH_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface pH versus Temperature", x = "Surface Temperature (C)", y = "Surface pH (NBS)") +
  theme_set(theme_bw())

```



Salinity and DO 
```{r}
ggplot(na.omit(ysidata), aes(Salinity_Surface, DO_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~SampleID) + 
  labs(title = "Surface Salinity versus Dissolved Oxygen in Mesocosm Tanks", x = "Surface Salinity (ppt)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())

```
Salinity and DO 
```{r}
ggplot(na.omit(ysidata), aes(Salinity_Surface, DO_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface Salinity versus Dissolved Oxygen", x = "Surface Salinity (ppt)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())

```


Temperature and Salinity 
```{r}
ggplot(na.omit(ysidata), aes(T_Surface, Salinity_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~SampleID) + 
  labs(title = "Surface Salinity versus Temperature in Mesocosm Tanks", x = "Surface Temperature (C)", y = "Surface Salinity (ppt)") +
  theme_set(theme_bw())

```

Overall Surface Salinity versus Temperature 

```{r}
ggplot(na.omit(ysidata), aes(T_Surface, Salinity_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface Salinity versus Temperature", x = "Surface Temperature (C)", y = "Surface Salinity (ppt)") +
  theme_set(theme_bw())

```

DO and pH by tank
```{r}
ggplot(na.omit(ysidata), aes(pH_Surface, DO_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~SampleID) + 
  labs(title = "Surface Dissolved Oxygen versus pH in Mesocosm Tanks", x = "Surface pH (NBS)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())


```


DO and pH 
```{r}
ggplot(na.omit(ysidata), aes(pH_Surface, DO_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface Dissolved Oxygen versus pH", x = "Surface pH (NBS)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())


```


DO and Temperature 

```{r}
ggplot(na.omit(ysidata), aes(T_Surface, DO_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~SampleID) + 
  labs(title = "Surface Dissolved Oxygen versus Temperature in Mesocosm Tanks", x = "Surface Temperature (C)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())

```
DO and Temperature 

```{r}
ggplot(na.omit(ysidata), aes(T_Surface, DO_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface Dissolved Oxygen versus Temperature", x = "Surface Temperature (C)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())

```

Salinity and pH

```{r}
ggplot(na.omit(ysidata), aes(pH_Surface, Salinity_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~SampleID) + 
  labs(title = "Surface Salinity versus pH in Mesocosm Tanks", x = "Surface pH (NBS)", y = "Surface Salinity (ppt)") +
  theme_set(theme_bw())

```
Salinity and pH

```{r}
ggplot(na.omit(ysidata), aes(pH_Surface, Salinity_Surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface Salinity versus pH", x = "Surface pH (NBS)", y = "Surface Salinity (ppt)") +
  theme_set(theme_bw())
```
CORRELATION MATRICES!!!
```{r}
### FOR SURFACE
surface_cor <- ysidata %>% 
  select(T_Surface, Salinity_Surface, pH_Surface, DO_Surface) %>%
  select(where(is.numeric)) %>% ## only select numeric variables
  drop_na()

cor(surface_cor) %>%
  corrplot()

### FOR DEEP
deep_cor <- ysidata %>% 
  select(T_Deep, Salinity_Deep, pH_Deep, DO_Deep) %>%
  select(where(is.numeric)) %>% ## only select numeric variables
  drop_na()

cor(deep_cor) %>% 
  corrplot()



```

CORRELATION STATS OF INTEREST 
```{r}
summary(lm(T_Surface~pH_Surface, data = ysidata))

summary(lm(DO_Surface~pH_Surface, data = ysidata))

summary(lm(T_Surface~DO_Surface, data = ysidata))
```

Box plots to compare mean values between different groups within the data set. 
Comparing pH between different tanks - high density versus  medium, low, and no density eel grass. 

```{r}

compare_sampleid <- list( c("3", "5"), c("3", "4"), 
                         c("3", "8"))

ggplot(na.omit(ysidata), aes(x=SampleID, y=T_Surface, fill=SampleID)) +
  geom_boxplot() + ## add boxplots
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_sampleid)


```


Try plotting Deep and Surface data sets on same graph for pH versus Temp

```{r}

ggplot(na.omit(ysidata), aes() + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface pH versus Temperature at Two Depths in the Mesocosm Tanks", x = "Temperature (C)", y = "pH (NBS)") +
  theme_set(theme_bw())

```


