---
title: "Daily.YSI.data"
author: "Kay"
date: "2023-06-15"
output: html_document
---

Load packages

``` {r}
require(pacman)
p_load(tidyverse,
       janitor,
       dplyr,
       ggpubr, 
       ggplot2,
       corrplot, 
       cowplot,
       readxl,
       here)
```


Use the 'here' function to map to directory 

``` {r}
here::i_am("YSIhandhelddata.Rmd") 
```


Import Excel file. Clean up data. 

``` {r}

ysidata <- read_excel(here("data/Daily_TankSpotChecks_2023.xlsx"))

ysidata <- ysidata[2:82,] %>% 
  rename("sample_id" = "Sample ID", "t_surface" = "T (C)", "t_deep" = "...4", "salinity_surface" = "Salinity (ppt)", "salinity_deep" = "...6", "pH_surface" = "pH (NBS)", "pH_deep" = "...8", "do_surface" = "DO (mg/L)", "do_deep" = "...10") %>%
  select(-Notes,-Initials,-`Calibration (y/n)`, - `Time Start(PSD)`, -`Time Stop(PSD)`, -`Flow Rate (L/min)`) 

ysidata <- ysidata %>%
  mutate_at(c('t_surface', 't_deep', 'salinity_surface', 'salinity_deep', 'pH_surface', 'pH_deep', 'do_surface','do_deep'), as.numeric) %>%
  mutate(density = case_when(sample_id == 1 ~ 'Medium', sample_id == 2 ~ 'High', sample_id == 3 ~ 'High', sample_id == 4 ~ 'Low', sample_id == 5 ~ 'Zero', sample_id == 6 ~ 'Low', sample_id == 7 ~ 'Zero', sample_id == 8 ~ 'Medium', sample_id == "Inlet" ~ "Inlet")) 

ysidata

```


Plot the relationship between Surface pH and Surface Temperature in all the tanks. 

``` {r}
ggplot(na.omit(ysidata), aes(t_surface, pH_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~density) + 
  ###facet_wrap(~sample_id) + 
  labs(title = "Surface pH versus Temperature in Mesocosm Tanks of Different Density Eelgrass", x = "Surface Temperature (C)", y = "Surface pH (NBS)") +
  theme_set(theme_bw())
```

Overall Surface Temp versus pH graph

```{r}
ggplot(na.omit(ysidata), aes(t_surface, pH_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface pH versus Temperature", x = "Surface Temperature (C)", y = "Surface pH (NBS)") +
  theme_set(theme_bw())
```

Salinity and DO

```{r}
ggplot(na.omit(ysidata), aes(salinity_surface, do_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~density) + 
  labs(title = "Surface Salinity versus Dissolved Oxygen in Mesocosm Tanks", x = "Surface Salinity (ppt)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())
```

Salinity and DO

```{r}
ggplot(na.omit(ysidata), aes(salinity_surface, do_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface Salinity versus Dissolved Oxygen", x = "Surface Salinity (ppt)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())
```

Temperature and Salinity 

```{r}
ggplot(na.omit(ysidata), aes(t_surface, salinity_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~density) + 
  labs(title = "Surface Salinity versus Temperature in Mesocosm Tanks", x = "Surface Temperature (C)", y = "Surface Salinity (ppt)") +
  theme_set(theme_bw())
```

Overall Surface Salinity versus Temperature 

```{r}
ggplot(na.omit(ysidata), aes(t_surface, salinity_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface Salinity versus Temperature", x = "Surface Temperature (C)", y = "Surface Salinity (ppt)") +
  theme_set(theme_bw())
```

DO and pH by tank

```{r}
ggplot(na.omit(ysidata), aes(pH_surface, do_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~density) + 
  labs(title = "Surface Dissolved Oxygen versus pH in Mesocosm Tanks", x = "Surface pH (NBS)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())
```

DO and pH 

```{r}
ggplot(na.omit(ysidata), aes(pH_surface, do_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface Dissolved Oxygen versus pH", x = "Surface pH (NBS)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())
```

DO and Temperature 

```{r}
ggplot(na.omit(ysidata), aes(t_surface, do_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~density) + 
  labs(title = "Surface Dissolved Oxygen versus Temperature in Mesocosm Tanks", x = "Surface Temperature (C)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())
```

DO and Temperature 

```{r}
ggplot(na.omit(ysidata), aes(t_surface, do_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface Dissolved Oxygen versus Temperature", x = "Surface Temperature (C)", y = "Surface Dissolved Oxygen (mg/L)") +
  theme_set(theme_bw())
```

Salinity and pH

```{r}
ggplot(na.omit(ysidata), aes(pH_surface, salinity_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  facet_wrap(~density) + 
  labs(title = "Surface Salinity versus pH in Mesocosm Tanks", x = "Surface pH (NBS)", y = "Surface Salinity (ppt)") +
  theme_set(theme_bw())
```

Salinity and pH

```{r}
ggplot(na.omit(ysidata), aes(pH_surface, salinity_surface)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Surface Salinity versus pH", x = "Surface pH (NBS)", y = "Surface Salinity (ppt)") +
  theme_set(theme_bw())
```


CORRELATION MATRICES!!!

```{r}
### FOR SURFACE
surface_cor <- ysidata %>% 
  select(t_surface, salinity_surface, pH_surface, do_surface) %>%
  select(where(is.numeric)) %>% ## only select numeric variables
  drop_na()

cor(surface_cor) %>%
  corrplot()

### FOR DEEP
deep_cor <- ysidata %>% 
  select(t_deep, salinity_deep, pH_deep, do_deep) %>%
  select(where(is.numeric)) %>% ## only select numeric variables
  drop_na()

cor(deep_cor) %>% 
  corrplot()
```

CORRELATION STATS OF INTEREST 

```{r}
summary(lm(t_surface~pH_surface, data = ysidata))

summary(lm(do_surface~pH_surface, data = ysidata))

summary(lm(t_surface~do_surface, data = ysidata))
```

Box plots to compare mean values between different groups within the data set. 
Comparing pH between different tanks - high density versus  medium, low, and no density eel grass. 

```{r}
compare_density <- list(c("Medium", "Low"), c("Low", "High"), 
                        c("High", "Medium"), c("High", "Zero"), 
                        c("Zero", "Medium"), c("Zero","Low"))

ggplot(na.omit(ysidata), aes(x=density, y=t_surface, fill = density)) +
  labs(title = "Surface Temperature in Mesocosm Tanks", x = "Eelgrass Density", y = "Surface Temperature (C)") + 
  geom_boxplot() + ## add boxplots
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_density)

### Figure out how to reorder densities  
```

pH Boxplots 

```{r}
ggplot(na.omit(ysidata), aes(x=density, y=pH_surface, fill = density)) +
  labs(title = "Surface pH in Mesocosm Tanks", x = "Eelgrass Density", y = "Surface pH (NBS)") + 
  geom_boxplot() + ## add boxplots
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_density)
```

Dissolved Oxygen Surface boxplot 

```{r}
ggplot(na.omit(ysidata), aes(x=density, y=do_surface, fill = density)) +
  labs(title = "Surface Dissolved Oxygen in Mesocosm Tanks", x = "Eelgrass Density", y = "Surface Dissolved Oxygen (mg/L)") + 
  geom_boxplot() + ## add boxplots
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_density)

```

Surface Salinity Boxplots 

```{r}
ggplot(na.omit(ysidata), aes(x=density, y=salinity_surface, fill = density)) +
  labs(title = "Surface Salinity in Mesocosm Tanks", x = "Eelgrass Density", y = "Surface Salinity (ppt)") + 
  geom_boxplot() + ## add boxplots
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_density)
```

DEEP pH Boxplot 

```{r}
ggplot(na.omit(ysidata), aes(x=density, y=pH_deep, fill = density)) +
  labs(title = "Deep pH in Mesocosm Tanks", x = "Eelgrass Density", y = "Deep pH (NBS)") +     geom_boxplot() + ## add boxplots
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_density)
```

Deep temp boxplot

```{r}
ggplot(na.omit(ysidata), aes(x=density, y= t_deep, fill = density)) +
  labs(title = "Deep Temperature in Mesocosm Tanks", x = "Eelgrass Density", y = "Deep Temperature (C)") + 
  geom_boxplot() + ## add boxplots
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_density)
```

Deep DO Boxplots

```{r}
ggplot(na.omit(ysidata), aes(x=density, y=t_surface, fill = density)) +
  labs(title = "Deep Dissolved Oxygen in Mesocosm Tanks", x = "Eelgrass Density", y = "Dissolved Oxygen (mg/L)") + 
  geom_boxplot() + ## add boxplots
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_density)
```

Deep Salinity boxplots 

```{r}
ggplot(na.omit(ysidata), aes(x=density, y=salinity_deep, fill = density)) +
  labs(title = "Deep Salinity in Mesocosm Tanks", x = "Eelgrass Density", y = "Deep Salinity (ppt)") + 
  geom_boxplot() + ## add boxplots
  stat_compare_means(aes(label = ..p.signif..), comparisons = compare_density)

```

Use pivot_longer to combine surface and deep columns and create new column 'depth'.Try plotting Deep and Surface data sets on same graph for pH versus depth 

```{r}
###Found this here: https://stackoverflow.com/questions/59253987/parallel-pivot-longer-of-two-sets-of-columns
## I don't know how to get what I want, so time to take the lazy way out
ysidata %>% 
  filter(!is.na(sample_id)) %>% 
  select(-sample_id) %>% 
  pivot_longer(-Date, 
               names_to = c("Var", ".value"), 
               names_sep = "_")


## Create two dataframes, one for surface, one for deep
df_surface <- ysidata %>% 
  filter(!is.na(sample_id)) %>% 
  select(-sample_id) %>%
  select(Date, density, contains("_surface")) %>% ## Select just surface columns + metadata
  rename_with(~str_remove(., '_surface')) %>% ## Remove _surface so columns will match
  mutate(depth = "surface")

df_deep <- ysidata %>% 
  filter(!is.na(sample_id)) %>% 
  select(-sample_id) %>%
  select(Date, density, contains("_deep")) %>%  
  rename_with(~str_remove(., '_deep')) %>% 
  mutate(depth = "deep")

df <- bind_rows(df_surface, df_deep)

## checking my work
ggplot(df, aes(pH, do, color = density)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = F) + 
  facet_wrap(~depth, nrow = 1)

ggplot(df, aes(density, do, fill = depth)) + 
  geom_boxplot()

```
